How to obtain the redshift distribution from probabilistic redshift estimates

### 2017-06-23 research/paper/draft-editing.tex

\documentclass[iop]{emulateapj}
%\documentclass[preprint]{aastex}
%\documentclass[12pt, onecolumn]{emulateapj}
%\documentstyle[aas2pp4,natbib209]{article}

@@

\begin{figure}
	\begin{center}
% 		\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{\textbf{Replace \texttt{daft} version with more polished \texttt{tikz} version.}}
%		\caption{This directed acyclic graph corresponds to a PGM for a hierarchical inference of $p(\vec{\theta}|\{\vec{d}_{i}\})$.  In this graph, all random variables are shown in circles, with observed random variables shown in shaded circles.  Relationships between variables are indicated by arrows from parameters to the variables distributed according to functions of them.  The box indicates that there are a number of copies of the relationships between boxed parameters, each independent of all others.  The hyperparameters $\vec{\theta}$ representing $n(z)$ are at the top.  Independently drawn from a function of the hyperparameters $\vec{\theta}$ are galaxy redshifts $\{z_{i}\}$ below.  The observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded circles, is determined by the redshifts above.}
	\label{fig:pgm}
	\end{center}
\end{figure}

@@

\begin{enumerate}
	\item Choose a true $n(z)$ that is a continuous function with known parameters $\vec{\theta}$.
	\item Sample the true $n(z)$ to create a catalog of $N$ true redshifts $z_{i}$.
	\item Choose a true intrinsic scatter $\sigma$, and sample Gaussians $p(z'_{i} | z_{i}, \sigma) = \mathcal{N}(z_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} | z'_{i}, \sigma) = \mathcal{N}(z'_{i}, \sigma)$.
	\item Choose a parametrization and the parameters $\vec{\phi}^{*}$ of the interim prior.
	\item Create interim posteriors $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) = p(\vec{d}_{i} | z_{i})\ p(z_{i} | \vec{\phi}^{*})$
\end{enumerate}

This method for deriving mock data is referred to as the fiducial case, and variations on it will refer directly to the steps that are altered.

@@

In this test, we choose a weakly featured true $n(z)$ of Fig. \ref{fig:smooth} with a smooth, unimodal shape.

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of a weakly featured true $n(z)$, probably SDSS interim prior]}
		\label{fig:featured}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of a highly featured, multimodal true $n(z)$]}
		\label{fig:featured}
	\end{center}
\end{figure}

@@

One major concern about photo-$z$s is the intrinsic scatter of point estimators, including those derived from photo-$z$ PDFs, that is observed to varying extents with every existing photo-$z$ algorithm and illustrated in Fig. \ref{fig:intscat}.

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with varying intrinsic scatter]}
		\label{fig:intscat}
	\end{center}
\end{figure}

To emulate intrinsic scatter, we modify the fiducial case to simply broaden the single Gaussian component of the likelihood.  To enforce self-consistency, the mean is a random variable drawn from a Gaussian distribution with the newly increased variance.

@@

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with template-like catastrophic outliers]}
		\label{fig:tempcatout}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with training-like catastrophic outliers]}
		\label{fig:traincatout}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of template-like (multimodal) interim prior]}
		\label{fig:tempintpr}
	\end{center}
\end{figure}

@@

An interim prior based on a training set may be biased toward low redshifts due to the dearth of distant galaxies with spectroscopic redshifts.  The interim prior shown in Fig. \ref{fig:trainintpr} is an emulation of an interim prior corresponding to a training set biased in this way.

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of training-like (low-$z$ favoring) interim prior]}
		\label{fig:trainintpr}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[moneyplot of error ellipses resulting from different $n(z)$ estimators]}
		\label{fig:money}
	\end{center}
\end{figure}

>>>>

\documentclass[iop]{emulateapj}
% \documentclass[preprint]{aastex}
% \documentclass[12pt, onecolumn]{emulateapj}
% \documentstyle[aas2pp4,natbib209]{article}

@@

\begin{figure}
	\begin{center}
 		\includegraphics[width=0.5\textwidth]{fig/pgm.png}
		\caption{This directed acyclic graph corresponds to a PGM for a hierarchical inference of $p(\vec{\theta}|\{\vec{d}_{i}\})$.  In this graph, all random variables are shown in circles, with observed random variables shown in shaded circles.  Relationships between variables are indicated by arrows from parameters to the variables distributed according to functions of them.  The box indicates that there are a number of copies of the relationships between boxed parameters, each independent of all others.  The hyperparameters $\vec{\theta}$ representing $n(z)$ are at the top.  Independently drawn from a function of the hyperparameters $\vec{\theta}$ are galaxy redshifts $\{z_{i}\}$ below.  The observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded circles, is determined by the redshifts above.}
	\label{fig:pgm}
	\end{center}
\end{figure}

@@

\begin{enumerate}
	\item \label{itm:true} Choose a true $n(z)$ that is a continuous function with known parameters $\vec{\theta}$.
	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of $N$ true redshifts $z_{i}$.
	\item \label{itm:scatter} Choose a true intrinsic scatter $\sigma$, and sample Gaussians $p(z'_{i} | z_{i}, \sigma) = \mathcal{N}(z_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} | z'_{i}, \sigma) = \mathcal{N}(z'_{i}, \sigma)$.
	\item \label{itm:interim} Choose a parametrization and the parameters $\vec{\phi}^{*}$ of the interim prior.
	\item \label{itm:posterior} Create interim posteriors $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) = p(\vec{d}_{i} | z_{i})\ p(z_{i} | \vec{\phi}^{*})$ in this parametrization.
\end{enumerate}

In all of the following validation tests, we use $10$ bins and $N = 10,000$ galaxies.  In Secs. \ref{sec:truth} and \ref{sec:likelihoods}, we use a flat interim prior.  This method for deriving mock data is referred to as the fiducial case, and variations on it will refer directly to the steps that are altered.

@@

In this test, we choose a weakly featured true $n(z)$ of Fig. \ref{fig:smooth} with a smooth, unimodal shape, based on the interim prior used for the SDSS DR8 photo-$z$ PDFs.  [check and include citation!]

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{fig/smooth_truth/scatter.png}\\
		\includegraphics[width=0.5\textwidth]{fig/smooth_truth/estimators.png}	
		\caption{All estimators perform well when the true $n(z)$ is well behaved, exhibiting significant deviation only when $n(z)$ is very small, as the sample size of true redshifts in that range will be small.  Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in solid lines, with a dotted line indicating zero probability.  Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
		\label{fig:featured}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{fig/fiducial/scatter.png}\\
		\includegraphics[width=0.5\textwidth]{fig/fiducial/estimators.png}		
		\caption{[This plot doesn't really show what I want because the intrinsic scatter is too low!  I used $\sigma=0.03$ because that's what has been quoted as what LSST, etc. needs, but that's not what comes out of photo-$z$ estimation methods before they impose aggressive cuts\dots]  Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in solid lines, with a dotted line indicating zero probability.  Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
		\label{fig:featured}
	\end{center}
\end{figure}

@@

One major concern about photo-$z$s is the intrinsic scatter of point estimators, including those derived from photo-$z$ PDFs, that is observed to varying extents with every existing photo-$z$ algorithm and illustrated in Fig. \ref{fig:intscat}.  To emulate intrinsic scatter, we modify the fiducial case to simply broaden the single Gaussian component of the likelihood.  To enforce self-consistency, the mean is a random variable drawn from a Gaussian distribution with the newly increased variance.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{fig/high_scatter/scatter.png}\\
		\includegraphics[width=0.5\textwidth]{fig/high_scatter/estimators.png}		
		\caption{As the intrinsic scatter increases, the discrepancy between estimators increases.  In particular, the stacked estimator and marginalized point estimators predict $\ln[n(z)]$ to be smoother than the truth, while the  [This would be a lot more compelling with more galaxies.  Also, the weird edge effects in the top panel are real, because the point estimator is the MAP, not the center of the Gaussian likelihood, and there's no requirement that the mean of the likelihood be within the true redshift range.]  Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in solid lines, with a dotted line indicating zero probability.  Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
		\label{fig:intscat}
	\end{center}
\end{figure}

As the intrinsic scatter increases, the discrepancy between estimators increases.

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{fig/template_outliers/scatter.png}\\
		\includegraphics[width=0.5\textwidth]{fig/template_outliers/estimators.png}		
		\caption{Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in solid lines, with a dotted line indicating zero probability.  Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
		\label{fig:tempcatout}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{fig/training_outliers/scatter.png}\\
		\includegraphics[width=0.5\textwidth]{fig/training_outliers/estimators.png}	
		\caption{[This has proven to be the most challenging test case to implement, and there's clearly still a bug in the function that makes the likelihoods.]}
		\label{fig:traincatout}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{fig/template_prior/scatter.png}\\
		\includegraphics[width=0.5\textwidth]{fig/template_prior/estimators.png}	
		\caption{[The case of a multimodal interim prior was a very compelling test in the previous version but somehow isn't anymore.]}
		\label{fig:tempintpr}
	\end{center}
\end{figure}

@@

An interim prior based on a training set may be biased toward low redshifts due to the dearth of distant galaxies with spectroscopic redshifts.  The interim prior shown in Fig. \ref{fig:trainintpr} is an emulation of an interim prior corresponding to a training set biased in this way.  We chose an interpolation of the interim prior used for the SDSS DR8 photo-$z$ PDFs.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{fig/training_prior/scatter.png}\\
		\includegraphics[width=0.5\textwidth]{fig/training_prior/estimators.png}	
		\caption{[The case of a low-$z$ favoring interim prior was a very compelling test in the previous version but somehow isn't anymore.]}
		\label{fig:trainintpr}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{money.png}
		\caption{[moneyplot of error ellipses resulting from different $n(z)$ estimators]}
		\label{fig:money}
	\end{center}
\end{figure}

### 2017-06-23 research/paper/draft.tex

In this test, we choose a weakly featured true $n(z)$ of Fig. \ref{fig:smooth} 
with a smooth, unimodal shape.

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of a weakly featured true $n(z)$, probably SDSS 
interim prior]}
		\label{fig:featured}
	\end{center}
\end{figure}

@@

\includegraphics[width=0.5\textwidth]{fig/fiducial/estimators.png}		
		\caption{[This plot doesn't really show what I want because the 
intrinsic scatter is too low!  I used $\sigma=0.03$ because that's what has 
been quoted as what LSST, etc. needs, but that's not what comes out of 
photo-$z$ estimation methods before they impose aggressive cuts.]  Top panel: 
The traditional MAP reduction of a photo-$z$ PDF against the true redshifts 
with a few rescaled photo-$z$ interim posteriors are overplotted in solid 
lines, with a dotted line indicating zero probability.  Bottom panel: Various 
estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a 
continuous function and under a binned parametrization.}
		\label{fig:featured}
	\end{center}
\end{figure}

@@

One major concern about photo-$z$s is the intrinsic scatter of point 
estimators, including those derived from photo-$z$ PDFs, that is observed to 
varying extents with every existing photo-$z$ algorithm and illustrated in Fig. 
\ref{fig:intscat}.

@@

\includegraphics[width=0.5\textwidth]{fig/high_scatter/estimators.png}		
		\caption{[This would be a lot more compelling with more 
galaxies.  Also, the weird edge effects in the top panel are real, because the 
point estimator is the MAP, not the center of the Gaussian likelihood, and 
there's no requirement that the mean of the likelihood be within the true 
redshift range.]  Top panel: The traditional MAP reduction of a photo-$z$ PDF 
against the true redshifts with a few rescaled photo-$z$ interim posteriors are 
overplotted in solid lines, with a dotted line indicating zero probability.  
Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the 
true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
		\label{fig:intscat}
	\end{center}
\end{figure}

To emulate intrinsic scatter, we modify the fiducial case to simply broaden the 
single Gaussian component of the likelihood.  To enforce self-consistency, the 
mean is a random variable drawn from a Gaussian distribution with the newly 
increased variance.

@@

		\caption{[]  Top panel: The traditional MAP reduction of a 
photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim 
posteriors are overplotted in solid lines, with a dotted line indicating zero 
probability.  Bottom panel: Various estimators of $\ln[n(z)]$, the interim 
prior, and the true $\ln[n(z)]$ as a continuous function and under a binned 
parametrization.}
		\label{fig:tempcatout}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with training-like 
catastrophic outliers]}
		\label{fig:traincatout}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of template-like (multimodal) interim prior]}
		\label{fig:tempintpr}
	\end{center}
\end{figure}

@@

An interim prior based on a training set may be biased toward low redshifts due 
to the dearth of distant galaxies with spectroscopic redshifts.  The interim 
prior shown in Fig. \ref{fig:trainintpr} is an emulation of an interim prior 
corresponding to a training set biased in this way.

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of training-like (low-$z$ favoring) interim 
prior]}
		\label{fig:trainintpr}
	\end{center}
\end{figure}

>>>>

In this test, we choose a weakly featured true $n(z)$ of Fig. \ref{fig:smooth} 
with a smooth, unimodal shape, based on the interim prior used for the SDSS DR8 
photo-$z$ PDFs.  [check and include citation!]

\includegraphics[width=0.5\textwidth]{fig/smooth_truth/scatter.png}\\

\includegraphics[width=0.5\textwidth]{fig/smooth_truth/estimators.png}	
		\caption{All estimators perform well when the true $n(z)$ is 
well behaved, exhibiting significant deviation only when $n(z)$ is very small, 
as the sample size of true redshifts in that range will be small.  Top panel: 
The traditional MAP reduction of a photo-$z$ PDF against the true redshifts 
with a few rescaled photo-$z$ interim posteriors are overplotted in solid 
lines, with a dotted line indicating zero probability.  Bottom panel: Various 
estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a 
continuous function and under a binned parametrization.}
		\label{fig:featured}
	\end{center}
\end{figure}

@@

\includegraphics[width=0.5\textwidth]{fig/fiducial/estimators.png}		
		\caption{[This plot doesn't really show what I want because the 
intrinsic scatter is too low!  I used $\sigma=0.03$ because that's what has 
been quoted as what LSST, etc. needs, but that's not what comes out of 
photo-$z$ estimation methods before they impose aggressive cuts\dots]  Top 
panel: The traditional MAP reduction of a photo-$z$ PDF against the true 
redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in 
solid lines, with a dotted line indicating zero probability.  Bottom panel: 
Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ 
as a continuous function and under a binned parametrization.}
		\label{fig:featured}
	\end{center}
\end{figure}

@@

One major concern about photo-$z$s is the intrinsic scatter of point 
estimators, including those derived from photo-$z$ PDFs, that is observed to 
varying extents with every existing photo-$z$ algorithm and illustrated in Fig. 
\ref{fig:intscat}.  To emulate intrinsic scatter, we modify the fiducial case 
to simply broaden the single Gaussian component of the likelihood.  To enforce 
self-consistency, the mean is a random variable drawn from a Gaussian 
distribution with the newly increased variance.

@@

\includegraphics[width=0.5\textwidth]{fig/high_scatter/estimators.png}		
		\caption{As the intrinsic scatter increases, the discrepancy 
between estimators increases.  In particular, the stacked estimator and 
marginalized point estimators predict $\ln[n(z)]$ to be smoother than the 
truth, while the  [This would be a lot more compelling with more galaxies.  
Also, the weird edge effects in the top panel are real, because the point 
estimator is the MAP, not the center of the Gaussian likelihood, and there's no 
requirement that the mean of the likelihood be within the true redshift range.] 
 Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true 
redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in 
solid lines, with a dotted line indicating zero probability.  Bottom panel: 
Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ 
as a continuous function and under a binned parametrization.}
		\label{fig:intscat}
	\end{center}
\end{figure}

As the intrinsic scatter increases, the discrepancy between estimators 
increases.

@@

		\caption{Top panel: The traditional MAP reduction of a 
photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim 
posteriors are overplotted in solid lines, with a dotted line indicating zero 
probability.  Bottom panel: Various estimators of $\ln[n(z)]$, the interim 
prior, and the true $\ln[n(z)]$ as a continuous function and under a binned 
parametrization.}
		\label{fig:tempcatout}
	\end{center}
\end{figure}

@@

\includegraphics[width=0.5\textwidth]{fig/training_outliers/scatter.png}\\

\includegraphics[width=0.5\textwidth]{fig/training_outliers/estimators.png}	
		\caption{[This has proven to be the most challenging test case 
to implement, and there's clearly still a bug in the function that makes the 
likelihoods.]}
		\label{fig:traincatout}
	\end{center}
\end{figure}

@@

\includegraphics[width=0.5\textwidth]{fig/template_prior/scatter.png}\\

\includegraphics[width=0.5\textwidth]{fig/template_prior/estimators.png}	
		\caption{[The case of a multimodal interim prior was a very 
compelling test in the previous version but somehow isn't anymore.]}
		\label{fig:tempintpr}
	\end{center}
\end{figure}

@@

An interim prior based on a training set may be biased toward low redshifts due 
to the dearth of distant galaxies with spectroscopic redshifts.  The interim 
prior shown in Fig. \ref{fig:trainintpr} is an emulation of an interim prior 
corresponding to a training set biased in this way.  We chose an interpolation 
of the interim prior used for the SDSS DR8 photo-$z$ PDFs.

\includegraphics[width=0.5\textwidth]{fig/training_prior/scatter.png}\\

\includegraphics[width=0.5\textwidth]{fig/training_prior/estimators.png}	
		\caption{[The case of a low-$z$ favoring interim prior was a 
very compelling test in the previous version but somehow isn't anymore.]}
		\label{fig:trainintpr}
	\end{center}
\end{figure}
